---
title: "Introduction to Rbitcoin"
author: "Jan Gorecki"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Introduction to Rbitcoin}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

## Introduction

Rbitcoin package can be used to create end-to-end trading engine in R.  
The goal of the vignettes is to present a possible workflow based on the functions available in the package. Additionally few other handy utilies are also presented.
```{r init, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", dev="svg")
suppressPackageStartupMessages(library(Rbitcoin))
Sys.setenv(TZ="UTC") # plot always in UTC
options("Rbitcoin.antiddos.sec"=5) # faster build vignette
```

There is also simple shinyApp as GUI for Rbitcoin package [shinyBTC](https://github.com/jangorecki/shinyBTC) which can be complementary to this vignette.  
```{r shinybtc, eval=FALSE}
shiny::runGitHub("jangorecki/shinyBTC")
```

## Market API

The core functionality of Rbitcoin is to communicate with cryptocurrency exchanges API directly from R, and to unify the structure of market API response across the different markets.  
Lets see the full process.

0. We will start by picking the market and currency pairs on which we will operate
```{r market_currency_pair}
market <- "kraken"
currency_pair <- c("BTC","EUR")
```

### Public API call

Public API calls do not require any authentication and you can query it without having an account on market.

1. At the beginning we might be interested to see top level price data using *ticker* API method
```{r ticker_api}
ticker <- market.api.process(market, currency_pair, "ticker")
ticker
```

2. a. Then we may want to lookup recent *trades*
```{r trades_api}
trades <- market.api.process(market,currency_pair,"trades")
trades[["trades"]][,tail(.SD,10)] # print only last 10 trades
```

2. b. Above call gives us a raw trades data which might be not so meaningful, we can visualize trades to see a little bit more
```{r trades_plot, fig.width = 7, fig.height = 5}
rbtc.plot(trades)
```

3. a. So far good but we want to confirm the *order book* is depth enough to fulfill our order
```{r order_book_api, fig.width = 7, fig.height = 5}
order_book <- market.api.process(market,currency_pair,"order_book")
rbtc.plot(order_book)
```

3. b. If we are going to buy we might want to check the top of *asks* from order book
```{r order_book_print}
order_book[["asks"]][,head(.SD,10)] # print only first 10 asks
```

### Private API call

Private API calls requires authentication, user need to have an account on market and generate API keys (`key` and `secret` param pair).  
Below examples will not be evaluated due to missing `key` and `secret` parameters in the vignette.

4. Query your *wallet* (account balance) information from market
```{r wallet_api, eval=FALSE}
wallet <- market.api.process(market, action = "wallet", key = "", secret = "")
wallet[["wallet"]] # print currencies and their amount in the wallet
```

5. *Place limit order* on the market (*buy* amount of *0.15* BTC at *500* EUR price per 1 BTC)
```{r place_limit_order_api, eval=FALSE}
place_limit_order <- market.api.process(market, currency_pair, action = "place_limit_order", 
                                        req = list(type = "buy",
                                                   price = 500,
                                                   amount = 0.15)
                                        key = "", secret = "")
```

6. Fetch currently *open orders* on the market
```{r open_orders_api, eval=FALSE}
open_orders <- market.api.process(market, action = "open_orders", key = "", secret = "")
```

7. *Cancel order* on market, `oid` must be provided (*oid* can be obtained using *open_orders* method)
```{r cancel_order_api, eval=FALSE}
cancel_order <- market.api.process(market, action = "cancel_order", 
                                   req = list(oid = "")
                                   key = "", secret = "")
```

### Market API notes

To avoid ban on API interface caused by a sequence of API calls, the `market.api.process` (and any other function which query over web) will perform antiddos procedure behind the scene (to customize see `?antiddos`).  
Because of the `market.api.process` function do post-process and pre-process of API calls to common structure across markets, the function is limited to defined markets and currency pairs in the dictionary (`?api.dict`). User can extend built-in dictionary for new markets (`?query.dict`) or new currency pairs or new API methods (`?api.dict`).  
In case if user wants to use currency_pair/method not defined in the dictionary it is possible using `market.api.query` function, this function can handle any currency pair and any method but it requires appropriate format of method arguments and it will not post-process API response to common structure.  
Built-in dictionary supports following market and currency pairs
```{r api.dict}
api.dict <- getOption("Rbitcoin.api.dict")
api.dict[!is.na(base), .(market, currency_pair = paste0(base,quote))][,unique(.SD)]
```

## Utilities

### to/from BTC

Simple conversion of fiat currencies to BTC (based on **blockchain.info** rates)
```{r to_from_api}
fromBTC(1) # current BTCUSD price
toBTC(150, "GBP") # convert 150 GBP to BTC
```

### Blockchain API

Query blockchain.info API
```{r blockchain_api}
# some first wallets btc address details
addr <- blockchain.api.process('1A1zP1eP5QGefi2DMPTfTL5SLmv7DivfNa')
str(addr)
# Rbitcoin donations btc address total received
blockchain.api.process('15Mb2QcgF3XDMeVn6M7oCG6CQLw4mkedDi')[["total_received"]]
# some transaction details
tx <- blockchain.api.process('e5c4de1c70cb6d60db53410e871e9cab6a0ba75404360bf4cda1b993e58d45f8')
str(tx, max.level=1)
```
If you need to query any other data from blockchain API you can use `blockchain.api.query` function which allows any methods but do not post-process response.

### Wallet Manager

Wallet manager is quite complex function to track your cryptocurrency balances (and it's values at a time) on different markets/accounts/addresses.  
By default function do not archive it's results, this must be setup using `archive_write=TRUE` args but user should be aware that sensitive data (balance and its sources **but not api keys**) will be archived locally in working directory as `wallet_archive.rds` file. Archive will allow `rbtc.plot` to plot historical assets balance over time. Historical balances might be also important for user's further analysis.  
Below is the example of sources definition and wallet manager execution.
```{r wallet_manager_exec, eval=FALSE}
# example market.sources
market.sources <- list(
  "john smith" = list(market='kraken', key='', secret=''),
  "jane smith" = list(market='kraken', key='', secret=''),
  "john smith" = list(market='btce', key='', secret=''),
  "jane smith" = list(market='btce', key='', secret='')
)
# example blockchain.sources
blockchain.sources <- list(
  "john smith" = list(address='')
)
# example manual.sources
manual.sources <- list(
  "john smith" = list(location='bitfinex', location_type='market',
                      currency=c('BTC','USD'), amount=c(0.4,0)),
  "john smith" = list(location='fidor', location_type='bank',
                      currency=c('EUR','USD'), amount=c(20,0)),
  "jane smith" = list(location='fidor', location_type='bank',
                      currency=c('EUR','GBP'), amount=c(10,105))
)
# execute
wallet_dt <- wallet_manager(
  market.sources = market.sources,
  blockchain.sources = blockchain.sources,
  manual.sources = manual.sources,
  value_currency = 'USD', # your target currency
  rate_priority = c('bitstamp','kraken','btce','bitmarket'), # value rates source priority
  archive_write = TRUE # by default FALSE, read ?wallet_manager
)
```
```{r populate_wallet_dt, echo=FALSE}
full_wallet_dt <- structure(
  list(wallet_id = c(1396310400L, 1396310400L, 1396310400L, 1396310400L, 1396310400L, 1396310400L, 1396310400L, 1396310400L, 1396310400L, 1396310400L, 1396310400L, 1396310400L,1396310400L, 1396310400L, 1396310400L, 1396310400L, 1396310400L, 1396310400L, 1398902400L, 1398902400L, 1398902400L, 1398902400L, 1398902400L, 1398902400L, 1398902400L, 1398902400L, 1398902400L, 1398902400L, 1398902400L, 1398902400L, 1398902400L, 1398902400L, 1398902400L, 1398902400L, 1398902400L, 1398902400L, 1401580800L, 1401580800L, 1401580800L, 1401580800L, 1401580800L, 1401580800L, 1401580800L, 1401580800L, 1401580800L, 1401580800L, 1401580800L, 1401580800L, 1401580800L, 1401580800L, 1401580800L, 1401580800L, 1401580800L, 1401580800L, 1404172800L, 1404172800L, 1404172800L, 1404172800L, 1404172800L, 1404172800L, 1404172800L, 1404172800L, 1404172800L, 1404172800L, 1404172800L, 1404172800L, 1404172800L, 1404172800L, 1404172800L, 1404172800L, 1404172800L, 1404172800L, 1406851200L, 1406851200L, 1406851200L, 1406851200L, 1406851200L, 1406851200L, 1406851200L, 1406851200L, 1406851200L, 1406851200L, 1406851200L, 1406851200L, 1406851200L, 1406851200L, 1406851200L, 1406851200L, 1406851200L, 1406851200L, 1409529600L, 1409529600L, 1409529600L, 1409529600L, 1409529600L, 1409529600L, 1409529600L, 1409529600L, 1409529600L, 1409529600L, 1409529600L, 1409529600L, 1409529600L, 1409529600L, 1409529600L, 1409529600L, 1409529600L, 1409529600L),
    currency = c("BTC", "BTC", "BTC", "BTC", "BTC", "EUR", "EUR", "EUR", "EUR", "GBP", "LTC", "LTC", "LTC", "USD", "USD", "USD", "USD", "USD", "BTC", "BTC", "BTC", "BTC", "BTC", "EUR", "EUR", "EUR", "EUR", "GBP", "LTC", "LTC", "LTC", "USD", "USD", "USD", "USD", "USD", "BTC", "BTC", "BTC", "BTC", "BTC", "EUR", "EUR", "EUR", "EUR", "GBP", "LTC", "LTC", "LTC", "USD", "USD", "USD", "USD", "USD", "BTC", "BTC", "BTC", "BTC", "BTC", "EUR", "EUR", "EUR", "EUR", "GBP", "LTC", "LTC", "LTC", "USD", "USD", "USD", "USD", "USD", "BTC", "BTC", "BTC", "BTC", "BTC", "EUR", "EUR", "EUR", "EUR", "GBP", "LTC", "LTC", "LTC", "USD", "USD", "USD", "USD", "USD", "BTC", "BTC", "BTC", "BTC", "BTC", "EUR", "EUR", "EUR", "EUR", "GBP", "LTC", "LTC", "LTC", "USD", "USD", "USD", "USD", "USD"),
    currency_type = c("crypto", "crypto", "crypto", "crypto", "crypto", "fiat", "fiat", "fiat", "fiat", "fiat", "crypto", "crypto", "crypto", "fiat", "fiat", "fiat", "fiat", "fiat", "crypto", "crypto", "crypto", "crypto", "crypto", "fiat", "fiat", "fiat", "fiat", "fiat", "crypto", "crypto", "crypto", "fiat", "fiat", "fiat", "fiat", "fiat", "crypto", "crypto", "crypto", "crypto", "crypto", "fiat", "fiat", "fiat", "fiat", "fiat", "crypto", "crypto", "crypto", "fiat", "fiat", "fiat", "fiat", "fiat", "crypto", "crypto", "crypto", "crypto", "crypto", "fiat", "fiat", "fiat", "fiat", "fiat", "crypto", "crypto", "crypto", "fiat", "fiat", "fiat", "fiat", "fiat", "crypto", "crypto", "crypto", "crypto", "crypto", "fiat", "fiat", "fiat", "fiat", "fiat", "crypto", "crypto", "crypto", "fiat", "fiat", "fiat", "fiat", "fiat", "crypto", "crypto", "crypto", "crypto", "crypto", "fiat", "fiat", "fiat", "fiat", "fiat", "crypto", "crypto", "crypto", "fiat", "fiat", "fiat", "fiat", "fiat"),
    auth = c("john smith", "john smith", "jane smith", "john smith", "john smith", "jane smith", "john smith", "john smith", "jane smith", "jane smith", "john smith", "jane smith", "jane smith", "john smith", "john smith", "john smith", "john smith", "jane smith", "john smith", "john smith", "jane smith", "john smith", "john smith", "jane smith", "john smith", "john smith", "jane smith", "jane smith", "john smith", "jane smith", "jane smith", "john smith", "john smith", "john smith", "john smith", "jane smith", "john smith", "john smith", "jane smith", "john smith", "john smith", "jane smith", "john smith", "john smith", "jane smith", "jane smith", "john smith", "jane smith", "jane smith", "john smith", "john smith", "john smith", "john smith", "jane smith", "john smith", "john smith", "jane smith", "john smith", "john smith", "jane smith", "john smith", "john smith", "jane smith", "jane smith", "john smith", "jane smith", "jane smith", "john smith", "john smith", "john smith", "john smith", "jane smith", "john smith", "john smith", "jane smith", "john smith", "john smith", "jane smith", "john smith", "john smith", "jane smith", "jane smith", "john smith", "jane smith", "jane smith", "john smith", "john smith", "john smith", "john smith", "jane smith", "john smith", "john smith", "jane smith", "john smith", "john smith", "jane smith", "john smith", "john smith", "jane smith", "jane smith", "john smith", "jane smith", "jane smith", "john smith", "john smith", "john smith", "john smith", "jane smith"),
    timestamp = structure(c(1396310400, 1396310400, 1396310400, 1396310400, 1396310400, 1396310400, 1396310400, 1396310400, 1396310400, 1396310400, 1396310400, 1396310400, 1396310400, 1396310400, 1396310400, 1396310400, 1396310400, 1396310400, 1398902400, 1398902400, 1398902400, 1398902400, 1398902400, 1398902400, 1398902400, 1398902400, 1398902400, 1398902400, 1398902400, 1398902400, 1398902400, 1398902400, 1398902400, 1398902400, 1398902400, 1398902400, 1401580800, 1401580800, 1401580800, 1401580800, 1401580800, 1401580800, 1401580800, 1401580800, 1401580800, 1401580800, 1401580800, 1401580800, 1401580800, 1401580800, 1401580800, 1401580800, 1401580800, 1401580800, 1404172800, 1404172800, 1404172800, 1404172800, 1404172800, 1404172800, 1404172800, 1404172800, 1404172800, 1404172800, 1404172800, 1404172800, 1404172800, 1404172800, 1404172800, 1404172800, 1404172800, 1404172800, 1406851200, 1406851200, 1406851200, 1406851200, 1406851200, 1406851200, 1406851200, 1406851200, 1406851200, 1406851200, 1406851200, 1406851200, 1406851200, 1406851200, 1406851200, 1406851200, 1406851200, 1406851200, 1409529600, 1409529600, 1409529600, 1409529600, 1409529600, 1409529600, 1409529600, 1409529600, 1409529600, 1409529600, 1409529600, 1409529600, 1409529600, 1409529600, 1409529600, 1409529600, 1409529600, 1409529600), class = c("POSIXct", "POSIXt"), tzone = "UTC"),
    location = c("bitfinex", "kraken", "kraken", "btce", "15Mb2QcgF3XDMeVn6M7oCG6CQLw4mkedDi", "fidor", "fidor", "kraken", "kraken", "fidor", "kraken", "kraken", "btce", "fidor", "bitfinex", "kraken", "btce", "btce", "bitfinex", "kraken", "kraken", "btce", "15Mb2QcgF3XDMeVn6M7oCG6CQLw4mkedDi", "fidor", "fidor", "kraken", "kraken", "fidor", "kraken", "kraken", "btce", "fidor", "bitfinex", "kraken", "btce", "btce", "bitfinex", "kraken", "kraken", "btce", "15Mb2QcgF3XDMeVn6M7oCG6CQLw4mkedDi", "fidor", "fidor", "kraken", "kraken", "fidor", "kraken", "kraken", "btce", "fidor", "bitfinex", "kraken", "btce", "btce", "bitfinex", "kraken", "kraken", "btce", "15Mb2QcgF3XDMeVn6M7oCG6CQLw4mkedDi", "fidor", "fidor", "kraken", "kraken", "fidor", "kraken", "kraken", "btce", "fidor", "bitfinex", "kraken", "btce", "btce", "bitfinex", "kraken", "kraken", "btce", "15Mb2QcgF3XDMeVn6M7oCG6CQLw4mkedDi", "fidor", "fidor", "kraken", "kraken", "fidor", "kraken", "kraken", "btce", "fidor", "bitfinex", "kraken", "btce", "btce", "bitfinex", "kraken", "kraken", "btce", "15Mb2QcgF3XDMeVn6M7oCG6CQLw4mkedDi", "fidor", "fidor", "kraken", "kraken", "fidor", "kraken", "kraken", "btce", "fidor", "bitfinex", "kraken", "btce", "btce"),
    location_type = c("market", "market", "market", "market", "blockchain", "bank", "bank", "market", "market", "bank", "market", "market", "market", "bank", "market", "market", "market", "market", "market", "market", "market", "market", "blockchain", "bank", "bank", "market", "market", "bank", "market", "market", "market", "bank", "market", "market", "market", "market", "market", "market", "market", "market", "blockchain", "bank", "bank", "market", "market", "bank", "market", "market", "market", "bank", "market", "market", "market", "market", "market", "market", "market", "market", "blockchain", "bank", "bank", "market", "market", "bank", "market", "market", "market", "bank", "market", "market", "market", "market", "market", "market", "market", "market", "blockchain", "bank", "bank", "market", "market", "bank", "market", "market", "market", "bank", "market", "market", "market", "market", "market", "market", "market", "market", "blockchain", "bank", "bank", "market", "market", "bank", "market", "market", "market", "bank", "market", "market", "market", "market"),
    amount = c(0, 0.15, 0, 0, 0.1, 190, 130, 40, 80, 110, 0, 4, 1, 275, 25, 10, 30, 30, 0, 0.15, 0, 0, 0.1, 10, 0, 170, 260, 0, 0, 4, 1, 0, 100, 110, 130, 205, 0.22, 0.6, 0.45, 0.3, 0.1, 10, 10, 0, 40, 15, 10, 8, 10.5, 0, 0, 5, 15, 105, 0.12, 0.1, 0.45, 0.3, 0.7, 10, 20, 0, 40, 105, 10, 0, 18.5, 0, 0, 5, 0, 5, 0, 0.1, 0.2, 0, 0.15, 10, 20, 0, 140, 105, 10, 5, 0, 0, 400, 5, 225, 160, 0.4, 0.1, 0.6, 0.4, 0.55, 10, 20, 0, 0, 105, 10, 5, 35, 0, 0, 5, 50, 15),
    value_currency = c("USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD", "USD"),
    value_rate = c(478.98, 478.98, 478.98, 478.98, 478.98, 1.380905, 1.380905, 1.380905, 1.380905, 1.674109, 13.24, 13.24, 13.24, 1, 1, 1, 1, 1, 460.97, 460.97, 460.97, 460.97, 460.97, 1.373171, 1.373171, 1.373171, 1.373171, 1.68398, 11.02, 11.02, 11.02, 1, 1, 1, 1, 1, 630.99, 630.99, 630.99, 630.99, 630.99, 1.359912, 1.359912, 1.359912, 1.359912, 1.691485, 11.04, 11.04, 11.04, 1, 1, 1, 1, 1, 638.25, 638.25, 638.25, 638.25, 638.25, 1.353611, 1.353611, 1.353611, 1.353611, 1.70729, 7.83, 7.83, 7.83, 1, 1, 1, 1, 1, 589.59, 589.59, 589.59, 589.59, 589.59, 1.331674, 1.331674, 1.331674, 1.331674, 1.669659, 7.67, 7.67, 7.67, 1, 1, 1, 1, 1, 474.54, 474.54, 474.54, 474.54, 474.54, 1.288716, 1.288716, 1.288716, 1.288716, 1.629786, 4.7, 4.7, 4.7, 1, 1, 1, 1, 1),
    value = c(0, 71.847, 0, 0, 47.898, 262.37195, 179.51765, 55.2362, 110.4724, 184.15199, 0, 52.96, 13.24, 275, 25, 10, 30, 30, 0, 69.1455, 0, 0, 46.097, 13.73171, 0, 233.43907, 357.02446, 0, 0, 44.08, 11.02, 0, 100, 110, 130, 205, 138.8178, 378.594, 283.9455, 189.297, 63.099, 13.59912, 13.59912, 0, 54.39648, 25.372275, 110.4, 88.32, 115.92, 0, 0, 5, 15, 105, 76.59, 63.825, 287.2125, 191.475, 446.775, 13.53611, 27.07222, 0, 54.14444, 179.26545, 78.3, 0, 144.855, 0, 0, 5, 0, 5, 0, 58.959, 117.918, 0, 88.4385, 13.31674, 26.63348, 0, 186.43436, 175.314195, 76.7, 38.35, 0, 0, 400, 5, 225, 160, 189.816, 47.454, 284.724, 189.816, 260.997, 12.88716, 25.77432, 0, 0, 171.12753, 47, 23.5, 164.5, 0, 0, 5, 50, 15)),
  .Names = c("wallet_id", "currency", "currency_type", "auth", "timestamp", "location", "location_type", "amount", "value_currency", "value_rate", "value"),
  row.names = c(NA, -108L),
  class = c("data.table", "data.frame"),
  sorted = c("wallet_id", "currency")
)
```

Function gathers all the wallet balances from specified sources, calculates its values in specified *value_currency* and returns following structurem (*value_currency* can be also cryptocurrency).
```{r wallet_manager_recent_str, echo=c(2)}
wallet_dt <- full_wallet_dt[wallet_id==max(wallet_id)] # ECHO OFF recent wallet from populated data
str(wallet_dt)
```

Simple plot of only recent balances, last wallet manager check on example dummy data.
```{r wallet_manager_recent_plot, echo=c(1), fig.width = 7, fig.height = 7}
rbtc.plot(wallet_dt) # type="recent" due to `archive_read=FALSE` so: wallet_dt[,length(unique(wallet_id))]==1
wallet_dt <- full_wallet_dt # ECHO OFF full wallet archive from populated data
```

After multiple runs of `wallet_manager(..., archive_write=TRUE)` we may plot historical balances.  
Simple low-resolution (monthly, n=6) example of wallet manager dummy data over time.
```{r wallet_manager_archive_load, eval=FALSE}
# load archive only
wallet_dt <- wallet_manager(archive_write=FALSE, archive_read=TRUE)
```
```{r wallet_manager_archive_plot_value, eval=c(1), fig.width = 7, fig.height = 7}
rbtc.plot(wallet_dt) # type="value" due to `archive_read=TRUE`
# in case of poor legend scaling it might be better to export plot to file
svg("wallet_manager_value.svg")
rbtc.plot(wallet_dt)
dev.off()
```

Notice the graphs in this html are in *svg* format so are well zoomable.

## Rbitcoin notes

If you find this package useful or you made profit from using it please consider donation to Rbitcoin [15Mb2QcgF3XDMeVn6M7oCG6CQLw4mkedDi](https://blockchain.info/address/15Mb2QcgF3XDMeVn6M7oCG6CQLw4mkedDi).  
Part of the received donations will likely be transferred to **Rbitcoin** dependencies if and when they will setup BTC donation address.

#### Dependencies used

 * [data.table](http://cran.r-project.org/web/packages/data.table/index.html): any data manipulation, totally replaces data.frames
 * [RCurl](http://cran.r-project.org/web/packages/RCurl/index.html): connect to the market API interface over https
 * [digest](http://cran.r-project.org/web/packages/digest/index.html): authorize on market API private api calls
 * [jsonlite](http://cran.r-project.org/web/packages/jsonlite/index.html): decode market API response json string to R objects
 * [R](http://www.r-project.org/): R Core

#### Contribution

[Rbitcoin github](https://github.com/jangorecki/Rbitcoin):  

 * report issues
 * make feature request
 * make pull request
 
## License

[MIT license](http://opensource.org/licenses/MIT)
