---
title: "Introduction to Rbitcoin"
author: "Jan Gorecki"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to Rbitcoin}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

The goal of the vignettes is to present a possible workflow based on the functions available in the package. Additionally few other handy utilies are also presented.

There is also simple GUI app for Rbitcoin package [shinyBTC](https://github.com/jangorecki/shinyBTC) which can be complementary to this vignette.  
```{r shinybtc, eval=FALSE}
shiny::runGitHub("jangorecki/shinyBTC")
```

```{r init, echo=FALSE}
suppressPackageStartupMessages(library(Rbitcoin))
```

## Market API

The core functionality of Rbitcoin is to communicate with cryptocurrency exchanges API directly from R, and to unify the structure of market API response across the different markets.  
Lets see the full process.

0. We will start by picking the market and currency pairs on which we will operate
```{r market_currency_pair}
market <- "kraken"
currency_pair <- c("BTC","EUR")
```

### Public API call

Public API calls do not require any authentication and you can query it without having an account on market.

1. At the beginning we might be interested to see top level price data using `ticker` API method
```{r ticker_api}
ticker <- market.api.process(market, currency_pair, "ticker")
ticker
```

2. a. Then we may want to lookup recent `trades`
```{r trades_api}
trades <- market.api.process(market,currency_pair,"trades")
trades[["trades"]][,tail(.SD,10)] # print only last 10 trades
```

2. b. Above call gives us a raw trades data which might be not so meaningful, we can visualize trades to see a little bit more
```{r trades_plot, fig.width = 7, fig.height = 5}
rbtc.plot(trades)
```

3. a. So far good but we want to confirm the order book is depth enough to fulfill our order
```{r order_book_api, fig.width = 7, fig.height = 5}
order_book <- market.api.process(market,currency_pair,"order_book")
rbtc.plot(order_book)
```

3. b. If we are going to buy we might want to check the top of *asks* from order book
```{r order_book_print}
order_book[["asks"]][,head(.SD,10)] # print only first 10 asks
```

### Private API call

Private API calls requires authentication, user need to have an account on market and generate API keys (`key` and `secret` param pair).  
Below examples will not be evaluated due to missing `key` and `secret` parameters in the vignette.

4. Query your wallet (account balance) information from market
```{r wallet_api, eval=FALSE}
wallet <- market.api.process(market, action = "wallet", key = "", secret = "")
wallet[["wallet"]] # print currencies and their amount in the wallet
```

5. Place limit order on the market (buy amount of 0.15 BTC at 500 EUR price per 1 BTC)
```{r place_limit_order_api, eval=FALSE}
place_limit_order <- market.api.process(market, currency_pair, action = "place_limit_order", 
                                        req = list(type = "buy",
                                                   price = 500,
                                                   amount = 0.15)
                                        key = "", secret = "")
```

6. Fetch currently open orders on the market
```{r open_orders_api, eval=FALSE}
open_orders <- market.api.process(market, action = "open_orders", key = "", secret = "")
```

7. Cancel opened order on market, `oid` must be provided (`oid` can be obtained using `open_orders` method)
```{r cancel_order_api, eval=FALSE}
cancel_order <- market.api.process(market, action = "cancel_order", 
                                   req = list(oid = "")
                                   key = "", secret = "")
```

### Notes

To avoid ban on API interface caused by a sequence of API calls, the `market.api.process` (and any other function which query over web) will perform antiddos procedure behind the scene (to customize see `?antiddos`).  
Because of the `market.api.process` function do post-process and pre-process of API calls to common structure across markets, the function is limited to defined markets and currency pairs in the dictionary (`?api.dict`). User can extend built-in dictionary for new markets (`?query.dict`) or new currency pairs or new API methods (`?api.dict`).  
In case if user wants to use currency_pair/method not defined in the dictionary it is possible using `market.api.query` function, this function can handle any currency pair and any method but it requires appropriate format of method arguments and it will not post-process API response to common structure.  
Built-in dictionary supports following market and currency pairs
```{r api.dict}
api.dict <- getOption("Rbitcoin.api.dict")
api.dict[!is.na(base), .(market, `currency pair` = paste0(base,quote))][,unique(.SD)]
```

## Utilities

### Blockchain API

Simple conversion of fiat currencies to BTC (based on blockchain.info rates)
```{r to_from_api}
fromBTC(1) # current BTCUSD price
toBTC(150, "GBP") # convert 150 GBP to BTC
```

Query blockchain.info API
```{r blockchain_api}
# some first wallets btc address details
addr <- blockchain.api.process('1A1zP1eP5QGefi2DMPTfTL5SLmv7DivfNa')
str(addr)
# some first wallets btc address final balance
blockchain.api.process('1A1zP1eP5QGefi2DMPTfTL5SLmv7DivfNa')[["final_balance"]]
# some transaction details
tx <- blockchain.api.process('e5c4de1c70cb6d60db53410e871e9cab6a0ba75404360bf4cda1b993e58d45f8')
str(tx, max.level=1)
```
If you need to query any other data from blockchain API you can use `blockchain.api.query` function which allows any methods but do not post-process response.

### Wallet Manager

Wallet manager is quite complex function to track your cryptocurrency balances in different markets/accounts.  
By default function do not archive it's results, this can be setup using `archive_write=TRUE` args but user should be aware that sensitive data will be archived localy in working directory as `wallet_archive.rds` file.  
Archive will allow `rbtc.plot` to plot historical cryptocurrency balances. Historical balances might be also important for user's further analysis.
```{r wallet_manager_api, eval=FALSE}
# sources definition
market.sources <- list(
  "my_account1" = list(market='kraken', key='', secret=''),
  "my_account1" = list(market='btce', key='', secret=''),
  "my_account2" = list(market='btce', key='', secret='') # second account possible
)
# define wallets on blockchain
blockchain.sources <- list(
  "my_account1" = list(address=''),
  "my_account1" = list(address='')
)
# define wallets manually
manual.sources <- list(
  "my_account1" = list(location='transferring', currency='LTC', amount=0.8),
  "my_account1" = list(location='brainwallet', currenc ='BTC', amount=0.1)
)
# execute
wallet_dt <- wallet_manager(
  market.sources = market.sources,
  blockchain.sources = blockchain.sources,
  manual.sources = manual.sources,
  value_currency = 'GBP', # your target currency
  rate_priority = c('bitstamp','kraken','bitmarket','btce')
  archive_write = TRUE # by default FALSE, read ?wallet_manager
)
print(wallet_dt)
# after multiple runs of wallet_manager(archive_write=TRUE) we can do plot
wallet_dt <- wallet_manager(archive_write=FALSE, archive_read=TRUE)
rbtc.plot(wallet_dt)
```

## Reference

### Dependencies used

 * [RCurl](http://cran.r-project.org/web/packages/RCurl/index.html): connect to the market API interface over https
 * [digest](http://cran.r-project.org/web/packages/digest/index.html): authorize on market API private api calls
 * [jsonlite](http://cran.r-project.org/web/packages/jsonlite/index.html): decode market API response json string to R objects
 * [data.table](http://cran.r-project.org/web/packages/data.table/index.html): any data manipulation, totally replaces data.frames

### Missing something?

Report, make Feature Request or even better Pull Request to [Rbitcoin github](https://github.com/jangorecki/Rbitcoin).

[MIT license](http://opensource.org/licenses/MIT)
